\chapter{Discussion and Further Work}
\label{sec:discussion}

\section{Verification Model Performance}

As shown by tables \ref{tab:coastline-compress} and \ref{tab:coastline-compress-py}, the new C++ predictor greatly outperforms the old Python predictor in terms of execution time. A factor seven improvement is achieved for the predictor, resulting in an improvement factor between $2$ and $4$ depending on the absolute error limit. As the error limit increases, the execution time of the encoder module increases, taking up a greater percentage of the total execution time. This diminishes the improvement due to the new predictor.

The increase in execution time of the encoder for higher absolute error limits is likely due to the increased number of samples that are encoded using low-entropy codes. This is caused by lower entropy of quantizer indices for higher error limits. The codewords of high-entropy codes can be calculated on-the-fly, resulting in fast execution time. However, the \acrshort{V2V} codes used for low-entropy samples require codeword tables, adding additional lookup time. By use of specialized algorithms for binary tree lookups, additional improvements to execution time might be gained.

The verification model by \citeauthor{vorhaugHyperspectralImageCompression2024} stores all intermediate values for both the predictor and encoder modules. By use of the sampler class (\autoref{sec:implementation}), the new predictor allows selectively storing intermediates, reducing the memory requirements. Additionally, storing of most intermediate values of the encoder module has been made optional. Detailed analysis of the memory usage of the verification model has not been performed in this work. However, before modification, the compressor was not able to operate on full \acrshort{hypso}-2 images, $N_z=120$, $N_y=598$ and $N_x=1092$, on a machine with 64 GB of memory. With reduced memory requirements, peak allocated memory for \acrshort{hypso}-2 images ranges between 30 and 40 GB. This is still a large amount of memory, but is low enough that successful evaluation of the delayed weight update scheme could be performed.

For use in the \acrshort{hypso} mission, compression is performed using the on-board hardware accelerator. This means that the verification model is not needed for compression during the acquisition of images. However, for on-ground decompression when the delayed weight update scheme is used, compliant software is required. This prompts the need for improving the execution time of the verification model, as the execution times (\autoref{tab:coastline-decompress}) upwards of half an hour ($\sim2000$ seconds) are unacceptable. As shown by the rewritten predictor module, rewriting the hybrid encoder module to C++ can provide great performance improvements, and is therefor suggested. Additionally, specialized algorithms for lookup in \acrshort{V2V} codeword tables should be utilized.

The encoder module uses large arrays that are modified and accessed in-place during calculations. A similar scheme was used by the Python predictor, but is somewhat mitigated by the sampler class of the C++ predictor. However, values needed by complex data dependencies, such as sample representatives and weight updates, are still stored in large arrays. Moving large amounts of data in and out of the memory caches for access is slow, placing a large execution time penalty on the program. In order to remove most of the required memory accesses, and improve execution time, the use of specialized buffers should be employed. These buffers would be similar to those used by the hardware accelerator of \citeauthor{vorhaugHyperspectralImageCompression2024}, storing and providing only values that are needed for calculations. Additional memory can be saved by use of C++ type templates that select the required data type based the dynamic range of the image that is processed.

\section{Compression Performance}

As shown by \autoref{tab:coastline-reduction}, the reduction in compression performance due to delayed weight updates is minimal for the \acrshort{hypso}-2 coastline image set, at less than one percent. The performance penalty for the AVIRIS Yellowstone set is significantly larger, at over four percent for the selected error limits. This is likely due to higher correlation between image samples, such that the images are well suited for predictive compression. The higher compression ratios that are achieved for these images, as shown in figures \ref{fig:compression-performance-hypso} and \ref{fig:compression-performance-aviris}, support this.

The quality of the weight update scheme does not affect the quality of the reconstructed images. This is because the quantizer and prediction loop guarantee that image samples can be reconstructed with at most $m_z(t)$ of error (\autoref{sec:compression}). Weight updates affect the accuracy of the predicted sample values, and thus the entropy of the prediction residuals. This, in turn, affects the achievable compression performance. For images with high correlation between samples, such as the AVIRIS Yellowstone set, the quality of the weight vectors is of higher importance. This results in the greater performance penalty due to the delayed weight updates that is observed as compared to the \acrshort{hypso}-2 images.

Comparing the compression performance of this work with the results of \citeauthor{jiaRemovalFeedbackLoop2025} is challenging. The work does not mention in detail the configurations that are used for compression, other than reduced prediction mode and narrow local sums. Additionally, AVIRIS images are grouped into categories depending on image characteristics such as coastline and urban. However, the exact image groups are not mentioned. By averaging the results of \citeauthor{jiaRemovalFeedbackLoop2025} for AVIRIS images, a reduction of compression performance between one and four percent is given. These number lie in a similar range as those achieved for AVIRIS Yellowstone images of this work (\autoref{tab:coastline-reduction}).

Compression performance is greatly dependant on the selected configurations for the issue 2 algorithm. Furthermore, results depend on the applicable image sensor and characteristics of the image motive. The configuration used for the evaluations of this work is based on the extensive analysis by \citeauthor{blanesPerformanceImpactParameter2019} \cite{blanesPerformanceImpactParameter2019}. Additionally, a simple analysis is made here to find the most suited number of prediction bands for \acrshort{hypso}-2 images. This is due to the direct impact of this parameter on the area footprint of the hardware accelerator, as shown by \citeauthor{vorhaugHyperspectralImageCompression2024} \cite{vorhaugHyperspectralImageCompression2024}. Results using both ordinary and delayed weight updates correspond well to the value suggested by \citeauthor{blanesPerformanceImpactParameter2019}, as shown in \autoref{fig:prediction_bands}, with the optimal value being $P=2$ or $P=3$.

\autoref{tab:psnr-compared} shows how the \acrshort{PSNR} of reconstructed \acrshort{hypso}-2 images drops as the absolute error limit is increased, starting as 226 dB for $a_z=1$. As mentioned, the quality of the reconstructed images is not affected by the weight update scheme. Selecting the correct error limits is a tradeoff between the desired compression ratio and the required quality of reconstructed images. Some work has been done to analyse the effects of lossy compression on hyperspectral images, such as \cite{chen2019compression, garcia2020competitive, garciavilchez2010impact}, but the threshold for acceptable loss varies on the usecases of the images. Finding suitable error limits for use in the \acrshort{hypso} mission requires extensive analysis, and is outside the scope of this work.

As shown by the work by \citeauthor{bakkenHYPSO1CubeSatFirst2023}, images captured by the camera on-board the \acrshort{hypso}-1 satellite contain more noise in the outer spectral bands. Similar characteristics apply to the \acrshort{hypso}-2 camera. This means that outer bands are less interesting from a research perspective. Future work should explore the possibility of using band-dependant error limits to increase compression performance. This is a feature of the issue 2 standard which allows error limits to be specified per-band, allowing greater compression of the outer, less interesting bands of \acrshort{hypso} images. Currently, the hardware accelerator does not support this feature. If the technique provides improved compression performance for the applicable images, hardware support should be added.

\section{Verification Coverage}

Before testing on real hardware, the hardware accelerator was verified in simulation. A large set of combinations of different configurations exists for the implementation. Exhaustive testing, that is checking all possible combinations, is not feasible. For this reason, random test data and configurations is used for verification. This technique was also, successfully, utilized by \citeauthor{vorhaugHyperspectralImageCompression2024}. In order to provide confidence as to the correct functioning of the design, a large set of data should be used. To further improve confidence, coverage data can be collected, counting how the different possible configurations are used over the set of test data, as shown in \autoref{app:verification-coverage}. Although \citeauthor{vorhaugHyperspectralImageCompression2024} did use the same verification approach using random data, he did not collect coverage data. The coverage data gathered in this work provides improved confidence in the correct functioning of, not only the modified design, but also the accelerator as a whole as implemented by \citeauthor{vorhaugHyperspectralImageCompression2024}.

\section{RTL Results}

Resource utilization estimates from implementations for the two selected \acrshort{FPGA}s show that the modified design is able to keep the low area footprint, as compared to the original design by \citeauthor{vorhaugHyperspectralImageCompression2024}. For the Zynq ZU7EV, \acrshort{LUT} and \acrshort{FF} usage has increased from 5475 to 6349 and 2635 to 3480 respectively. The increase is mostly due to the extra pipeline registers required by the addition of the weight update pipeline and splitting of sample representative calculation. This is also shown by the number of \acrshort{BRAM}s, which is increased from $57.5$ to $79.5$. The other contributing factor is the weight refinement step, which requires the use of two additional adders for each of the weight vector components as shown in \autoref{fig:weights-pipeline-hw}. Interestingly, the number of \acrshort{DSP}s remains the same, in spite of the additional addition operations. Possibly, increased computational performance could be gained by ensuring costly operations are mapped to \acrshort{DSP} blocks. Detailed analysis of this has not been performed in this work.

Alleviation of the critical path in the weight update stage successfully increased the maximum achievable clock speed of the design. For the Zynq ZU7EV, the speed is increased from 141 MHz to 202 MHz for the entire compression core. In the original design by \citeauthor{vorhaugHyperspectralImageCompression2024}, the predictor stage was the limiting factor in terms of speed (141 MHz). The updated design is limited by the encoder module, with the predictor reaching a speed of 351 MHz in isolation. As explained in \autoref{sec:results}, the critical path of the predictor, after splitting stage D+9, is again in the final stage of the weight update calculation. This path runs through the weight refinement step, including the mentioned addition operations. Additional gains in speed could likely be gained by shifting the operations of the weight update pipeline to the left. That is, moving the addition of weight refinement to stage 3c, the increment of 1 in stage 3c to stage 3b, and the subtraction operator of stage 3b to stage 3a (\autoref{fig:weights-pipeline-hw}).

The estimated power consumption of design, for the Zynq ZU7EV, is 890 mW at a frequency of 202 MHz. This is somewhat higher than that of \citeauthor{vorhaugHyperspectralImageCompression2024} at 741 mW at 141 MHz. The difference is caused by two main factors: increased resource utilization and clock speed. As the number of components of a digital circuit increases, the number of transistors that must switch state at each clock edge is increased. Additionally, the higher clock frequency means that these transistors must switch faster. Both of these factors contribute to increased dynamic power consumption. Static power consumption is also increased as the number of components increases.

The work by \citeauthor{basconesRealTimeFPGAImplementation2022} shows that implementation of a faster hybrid encoder module is possible \cite{basconesRealTimeFPGAImplementation2022}. Their design reaches a clock speed of $\sim 300$ MHz on the XQRKU060 \acrshort{FPGA}, as opposed to 165 MHz reached by the encoder of \citeauthor{vorhaugHyperspectralImageCompression2024}. As shown in this work, the encoder reaches a somewhat higher frequency of 202 MHz on the Zynq ZU7EV. In addition to the predictor and encoder modules, the design comprises two additional modules in its processing pipeline: a reorder buffer and packer module (\autoref{fig:vorhaug-block}). As shown in this work, the \acrshort{BIP} to \acrshort{BIL} reorder buffer is able to match the frequency of 202 MHz. \citeauthor{vorhaugHyperspectralImageCompression2024} shows that the packer module can operate at a high frequency of 600 MHz on the Kintex XCKU040 \acrshort{FPGA}. These results show that future improvements to the design should focus on improving the encoder module.

As discussed in section \autoref{sec:previous-work}, the current design does not support run-time configuration. This is one of the key factors in keeping the area footprint low. For use on-board the \acrshort{hypso} missions, future work should aim at introducing run-time configuration for some of the key parameters. Easy selection of error limits would allow researchers to tune the loss in image quality on a per-image basis. Additionally, support for band-dependant error limits should be added. These additions should be made while minimizing the impact on resource utilization.

\section{Place in State-of-the-art}

\begin{table}[ht]
	\caption{Hardware utilization and throughput of hardware accelerators of the issue 2 standard. The predictor module is the smallest and fastest in open literature.}
	\label{tab:throughput-this-work}
	\centering
	\begin{tabular}{l r r r r r r}
		\hline
		\textbf{Implementation}                                                                                      & \textbf{FPGA} & \textbf{LUT} & \textbf{FF} & \textbf{DSP} & \textbf{BRAM} & \textbf{Throughput} \\
		\hline
		Barrios et al.\ \cite{barriosHardwareImplementationCCSDS2021}                                                & XCKU040       & 17185        & 11915       & 63           & 85            & 18~MS/s             \\
		B\'ascones et al.\ \cite{basconesRealTimeFPGAImplementation2022}                                             & XQRKU060      & 16458        & 15707       & 30           & 187.5         & 285~MS/s            \\
		Chatziantoniou et al.\ \cite{chatziantoniouScalableDataRateNearLossless2022a} 1 core                         & XCKU040       & 18861        & 21395       & -            & 104           & 285~MS/s            \\
		Chatziantoniou et al.\ \cite{chatziantoniouScalableDataRateNearLossless2022a} 5 core                         & XCKU040       & 67362        & 68987       & -            & 306           & 1375~MS/s           \\
		S\'anchez et al.\ \cite{sanchezReducingDataDependencies2022, barriosRemovingDataDependencies2024}, predictor & XCKU040       & 19840        & 12970       & 33           & 213           & 231~MS/s            \\
		Barrios et al.\ \cite{barriosRemovingDataDependencies2024}, predictor                                        & XCKU040       & 11095        & 11729       & 20           & 186.5         & 236~MS/s            \\
		Vorhaug et al.\ \cite{vorhaugHyperspectralImageCompression2024}                                              & XCKU040       & 5577         & 2635        & 4            & 57.5          & 110~MS/s            \\
		Vorhaug et al.\ \cite{vorhaugHyperspectralImageCompression2024}                                              & XCZU7EV       & 5475         & 2635        & 4            & 57.5          & 141~MS/s            \\
		Jia et al.\ \cite{jiaRemovalFeedbackLoop2025}, predictor                                                     & XCZU5EV       & 3995         & 5050        & 7            & 94            & 348~MS/s            \\
		This work                                                                                                    & XCZU7EV       & 6349         & 3480        & 4            & 79.5          & 202~MS/s            \\
		This work, predictor                                                                                         & XCZU7EV       & 2807         & 2244        & 4            & 69.5          & 351~MS/s            \\
		\hline
	\end{tabular}
\end{table}

\autoref{tab:throughput-this-work} summarizes achieved hardware resource utilization and throughput of issue 2 accelerators in the open literature, including the results of this work. The predictor of this work reaches a throughput of 351 MS/s, slightly outperforming that of \citeauthor{jiaRemovalFeedbackLoop2025}, whom proposed the delayed weight update scheme. Although not as fast as other implementations, the design by \citeauthor{vorhaugHyperspectralImageCompression2024} is considerable smaller than other implementations. This property is preserved in this work, resulting in both the smallest and fastest predictor of known implementations. Due to the limiting encoder module, the throughput of the entire compression core is not able to match that of \citeauthor{basconesRealTimeFPGAImplementation2022} and \citeauthor{chatziantoniouScalableDataRateNearLossless2022a}. However, future improvements should be able to alleviate this limitation.

In terms of estimated power consumption, with the exception of the accelerator of \citeauthor{vorhaugHyperspectralImageCompression2024}, the design of this work outperforms that of other works at 890 mW (\autoref{tab:power-this-work}). This is largely due to the low resource utilization. However, at the high clock frequency of 351 MHz achieved by the predictor module the power consumption of 1282 mW is beaten by the designs of \citeauthor{basconesRealTimeFPGAImplementation2022} and \citeauthor{chatziantoniouScalableDataRateNearLossless2022a}.

\begin{table}[ht]
	\caption{Power estimates of implementations in the open literature.}
	\label{tab:power-this-work}
	\centering
	\begin{tabular}{lrrr}
		\hline
		\textbf{Implementation}                                                                                      & \textbf{FPGA} & \textbf{Power} & \textbf{Throughput} \\
		\hline
		Barrios et al.\ \cite{barriosHardwareImplementationCCSDS2021}                                                & XCKU040       & 2480~mW        & 18~MS/s             \\
		B\'ascones et al.\ \cite{basconesRealTimeFPGAImplementation2022}                                             & XQRKU060      & 1210~mW        & 285~MS/s            \\
		Chatziantoniou et al.\ \cite{chatziantoniouScalableDataRateNearLossless2022a} 1 core                         & XCKU040       & 1090~mW        & 285~MS/s            \\
		Chatziantoniou et al.\ \cite{chatziantoniouScalableDataRateNearLossless2022a} 5 core                         & XCKU040       & 4222~mW        & 1375~MS/s           \\
		S\'anchez et al.\ \cite{sanchezReducingDataDependencies2022, barriosRemovingDataDependencies2024}, predictor & XCKU040       & -              & 231~MS/s            \\
		Barrios et al.\ \cite{barriosRemovingDataDependencies2024}, predictor                                        & XCKU040       & -              & 236~MS/s            \\
		Vorhaug et al.\ \cite{vorhaugHyperspectralImageCompression2024}                                              & XCKU040       & 640~mW         & 110~MS/s            \\
		Vorhaug et al.\ \cite{vorhaugHyperspectralImageCompression2024}                                              & XCZU7EV       & 741~mW         & 141~MS/s            \\
		Jia et al.\ \cite{jiaRemovalFeedbackLoop2025}, predictor                                                     & XCZU5EV       & -              & 348~MS/s            \\
		This work                                                                                                    & XCZU7EV       & 890~mW         & 202~MS/s            \\
		This work, predictor                                                                                         & XCZU7EV       & 1282mW         & 351~MS/s            \\
		\hline
	\end{tabular}
\end{table}

\section{CCSDS 123.0-B-2 Compliance}

As shown by \autoref{tab:compliance}, the accelerator developed by \citeauthor{vorhaugHyperspectralImageCompression2024} is restricted to a sub-set of the configurations of the issue 2 standard. As discussed in sections \ref{sec:state-of-the-art} and \ref{sec:previous-work}, most other implementations of the open literature also only support sub-sets of the standard. The modified accelerator, using delayed weight updates, still supports the same sub-set of configurations as the original design by \citeauthor{vorhaugHyperspectralImageCompression2024}. However, the use of delayed weight updates means that the accelerator is no longer compliant with the issue 2 standard. This means that compressed images cannot be decompressed using issue 2 compliant software tools, requiring the use of specialized software such as the decompressor developed in this work. A similar trade-off is made by the prequantization approach by \citeauthor{chatziantoniouScalableDataRateNearLossless2022a} and \citeauthor{valsesiaHighThroughputOnboardHyperspectral2019}, as well as the method of skipping weight updates proposed by \citeauthor{barriosRemovingDataDependencies2024}.

\section{On-board Testing}

The successful on-board compression, downlink and decompression of an image from the \acrshort{hypso}-2 satellite shows that the delayed weight update scheme is a viable solution for increased throughput of issue 2 accelerators. The updated design was run at a clock frequency of 92 MHz on the Zynq-7030, matching the frequency estimate given by static timing analysis. The original design by \citeauthor{vorhaugHyperspectralImageCompression2024} was tested on-board \acrshort{hypso}-1, and ran at a lower frequency of 50 MHz. Future improvements to the design should aim at reaching 100 MHz, which would allow the removal of the additional clocking wizard block (\autoref{app:block-design}), as the surrounding systems run at this frequency.

During integration for \acrshort{hypso}-2, an error was encountered in which the accelerator produced the wrong outputs. This turned out to be caused by the wrong selection of byte endianness on the output transmitted by the \acrshort{AXI} bus. The final image files, that are stored to the on-board SD-card use the big endian byte order. Because of this it was assumed that the same ordering should be used for transmission on the output data bus. This assumption proved wrong, and little endian had to be used. At some point in the on-board image processing, between retrieval from the compressor and storage, the byte order is switched.

Another caveat of the design was discovered during integration testing for the on-ground engineering model. For the first compression run after reprogramming the \acrshort{FPGA}, the accelerator produced the correct output. However, subsequent runs produced wrong outputs that were different every time. This suggests that an internal register expects an initial value for performing the correct calculations. The hardware design includes a reset signal, which would likely provide the correct initial value for the register if toggled. However, the \acrshort{hypso} \acrshort{OPU} does not toggle the reset signals of the loaded design between compression runs, causing the varying results. Localizing the problematic register has not been done in this work, as on-board tests could be carried out by reprogramming the \acrshort{FPGA} between compression runs.

The processing systems of the \acrshort{hypso}-1 and -2 satellites are not designed for on-the-fly compression. Images are buffered in \acrshort{RAM} before compression is initiated. This means that the throughput of the compression accelerator is not of high importance for these missions. The compressed bitstreams are stored on the on-board SD-card while awaiting downlink. The bottleneck of the downlink process, aside from waiting for contact with the ground station, is the transfer of data from the \acrshort{OPU} to the \acrshort{PC}. Here, a \acrfull{CAN} data bus with a transfer rate of 290 kb/s is used \cite{grotteOceanColorHyperspectral2022}.

For an image of size 156 MB, compressed at a ratio of 2.5 for a compressed bitstream size of 62.4 MB, a reduction in compression performance of $1\%$ means an additional 600 kB to be transferred over the \acrshort{CAN} bus. At the 290 kb/s data rate, this takes 16.6 seconds. Increasing the throughput of the compression accelerator from 50 MS/s to 92 MS/s reduces the compression time for the same image from 3.12 seconds to 1.70 seconds. This shows that, for the \acrshort{hypso}-1 and 2 satellites, the increased throughput gained by the delayed weight update scheme does not outweigh the overhead of transmitting the additional data of the downlink bottleneck. Future missions, such as \acrshort{hypso}-3, may replace the slow \acrshort{CAN} interface with an interface of significantly higher throughput \cite{grotteOceanColorHyperspectral2022}. This can be paired with a design which allows on-the-fly compression, increasing the significance of the accelerator throughput.

The \acrshort{hypso} \acrshort{OPU} is tailored for using an issue 1 compliant compression accelerator. Similarly to the issue 2 accelerator presented in this work and by \citeauthor{vorhaugHyperspectralImageCompression2024}, the accelerator outputs only the compressed bitstream body with accompanying encoder image tail. After compression, on-board software preprends the encoded issue 1 header config. This means that the images compressed on hardware using the issue 2 accelerator, both in-orbit and on-ground, were combined with this issue 1 header, rather than the correct issue 2 header. Because the header configuration of the hardware is known, and that decoding preoceeds in the reverse order, the images could still be decoded. For future integration, software for prepending the correct header should be written in order to provide compressed bitstreams of full issue 2 compliance.
